---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r set_global, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(fig.height=4)
```

## Loading and preprocessing the data
The data to be analyzed is included in the archive activity.zip.The full data set is loaded and a copy of it is created without the missing values = NA. Both datasets, with and without missing values, are needed to answer the following questions.

```{r Loading and Cleaning}
if(!file.exists("activity.zip")){
        stop("activity.zip file not present")
}
unzip("activity.zip")
All_Steps <- read.csv("activity.csv", colClasses = c("integer", "Date", "integer"))
OK_Steps <- na.omit(All_Steps)

```

As an example the first rows of the cleaned data frame are as follows:

```{r show rows}
OK_Steps[1:5,]

```


## What is mean total number of steps taken per day?

Using the OK_Steps that is not including NA values, the number of steps per day over the two months observation are showed in the following picture


```{r Steps_per_day}

Day_Steps <- with(OK_Steps,aggregate(steps, by = list(date), sum))
names(Day_Steps) <- c("date", "steps")
Mean_Steps <- as.integer(mean(Day_Steps$step))
Median_Steps <- as.integer(median(Day_Steps$step))

library(ggplot2)
g <- ggplot(Day_Steps, aes(date,steps)) + geom_bar(stat="identity")
g <- g + labs(title="Number of Steps per day")
g <- g + geom_hline(yintercept = c(Mean_Steps,Median_Steps), color = "blue", size = 1)
plot(g)

```

The picture is also showing in blue the median = `r Median_Steps` and mean = `r Mean_Steps` number of steps per day over the two months observation period. 




## What is the average daily activity pattern?
Moving to the 5 minutes data resolution, the activity pattern over the 2 months observation period, is reported in the following graph.

```{r Steps_5_min}
Mean_Step_5 <- as.integer(mean(OK_Steps$steps))
iMax_Step_5 <- which.max(OK_Steps$steps)
Max_Step <- OK_Steps[iMax_Step_5,]

g <- ggplot(OK_Steps, aes(date,steps)) + geom_line()
g <- g + labs(title = "Number of Steps per 5 minutes")
g <- g + geom_hline(yintercept = Mean_Step_5, color = "blue", size = 1)
g <- g + geom_point(data=Max_Step, aes(date,steps), color = "red", size = 4)
plot(g)
```

The picture is also showing in blue the mean = `r Mean_Step_5` number of steps every 5 minutes over the two months observation period. 

Finally the maximum number of steps made in 5 minutes period is equal to `r Max_Step$step` and occurred on the `r Max_Step$date` at time interval `r Max_Step$interval`.




## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
